{\bf Intrusion Detection through Provenance-Based Histograms}

### 2013-11-24 writeup/paper.tex
typoe in construction

\subsection{Histograms by Name}
Currently, our system compute statistics based on node type, but not node name. For instance, When we compute the outgoing (i.e. "forked") processes, we simply count them. Thus, if a program normally forks something benign, like say \texttt{cp}, but an exploited version forks \texttt{rm}, the raw counts may not reveal this (a call to \texttt{/bin/sh}, however, should impact the centrality measures). One direction to attempt to rectify this problem lies in generating more specific histograms. Instead of having a model for "number of processes forked", we could generate one model per type of process forked, e.g. "number of \texttt{cp} processes forked". This more specific data might enable us to catch intrusions in larger programs that already modify a lot of files and fork a lot of processes.
\subsection{ENV and ARGV}
Currently we ignore the environment and argument information stored in the provenance graph. By generating histograms based on this information, we could catch subtler intrusions/exploits that rely on modified environment information. Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
\subsection{Parameters}
Currently, constructing the kernel density estimates requires a few parameter decisions - namely the number/size of the bins for our histograms, as well as the bandwidth parameter for construction the KDE. We, for the purposes of this experiment, modified these parameters manually until they described the data well, but we could computer optimal values programmatically, or test out multiple values. Additionally, we demonstrated that odd behavior has a ``low" KDE, but we did not define what ``low" means. For this system to be deployed more widely, we will need to define or compute what thresholds signify a potential intrusion or exploit.
\subsection{Online Intrusion Detection}
Currently, this system runs after-the-fact. We can identify intrusions that occurred on a machine by analyzing the data after it has been lifted from the target machine. Because PASS collect data in real time, however, our approach can be modified to implement online intrusion detection. To achieve this goal, we would need to only use metrics that can be updated without recomputing the entire statistic. In particular, some of the centrality measures do not meet this criterion. We would also need to add additional daemons (or modify the existing PASS daemons) to run the appropriate statistic gathering and model generating code. Another extension is to apply progressive versions of our centrality metrics, including a progressive form of PageRank which computes the importance of a node online.
\subsection{Local vs. Global}
In our experiments, we looked at provenance graphs in their entirety. However, provenance graphs are meant to be long-lived, and an intrusion that affected only a small portion of the graph will have negligible effects on the global structure of the graph. To resolve this issue, a possible future extension to our work is to cluster the nodes based on time \cite{clustering} and then only examine the most ``recent" nodes when applying our histogram technique.
\section{Conclusion}
We have demonstrated that provenance data (in this case, data generated by PASS) can be used to identify anomalies in program behavior, which often signal a security intrusion or exploit. Using provenance-based histograms, we can construct models of the regular usage patterns for a particular program. When the intrusion or exploit results in behavior deviating from these models, its provenance allows us to flag it as a potential intrusion. Replaced binaries that acted significantly differently were especially easy to detect. Histograms on simple provenance statistics are good for identifying simple anomalies, but more complex statistics (see future work) may enable us to detect subtler intrusions.

>>>>

\subsection{Histograms by Name}
Currently, our system compute statistics based on node type, but not node name. For instance, When we compute the outgoing (i.e. "forked") processes, we simply count them. Thus, if a program normally forks something benign, like say \texttt{cp}, but an exploited version forks \texttt{rm}, the raw counts may not reveal this (a call to \texttt{/bin/sh}, however, should impact the centrality measures). One direction to attempt to rectify this problem lies in generating more specific histograms. Instead of having a model for "number of processes forked", we could generate one model per type of process forked, e.g. "number of \texttt{cp} processes forked". This more specific data might enable us to catch intrusions in larger programs that already modify a lot of files and fork a lot of processes.
\subsection{ENV and ARGV}
Currently we ignore the environment and argument information stored in the provenance graph. By generating histograms based on this information, we could catch subtler intrusions/exploits that rely on modified environment information. Additionally, if an exploit relies on a particular argument or argument patterns, this information might reflected in the ARGV information.
\subsection{Parameters}
Currently, constructing the kernel density estimates requires a few parameter decisions - namely the number/size of the bins for our histograms, as well as the bandwidth parameter for constructing the KDE. We, for the purposes of this experiment, modified these parameters manually until they described the data well, but we could computer optimal values programmatically, or test out multiple values. Additionally, we demonstrated that odd behavior has a ``low" KDE, but we did not define what ``low" means. For this system to be deployed more widely, we will need to define or compute what thresholds signify a potential intrusion or exploit.
\subsection{Online Intrusion Detection}
Currently, this system runs after-the-fact. We can identify intrusions that occurred on a machine by analyzing the data after it has been lifted from the target machine. Because PASS collect data in real time, however, our approach can be modified to implement online intrusion detection. To achieve this goal, we would need to only use metrics that can be updated without recomputing the entire statistic. In particular, some of the centrality measures do not meet this criterion. We would also need to add additional daemons (or modify the existing PASS daemons) to run the appropriate statistic gathering and model generating code. Another extension is to apply progressive versions of our centrality metrics, including a progressive form of PageRank which computes the importance of a node online.
\subsection{Local vs. Global}
In our experiments, we looked at provenance graphs in their entirety. However, provenance graphs are meant to be long-lived, and an intrusion that affected only a small portion of the graph will have negligible effects on the global structure of the graph. To resolve this issue, a possible future extension to our work is to cluster the nodes based on time \cite{clustering} and then only examine the most ``recent" nodes when applying our histogram technique.
\section{Conclusion}
We have demonstrated that provenance data (in this case, data generated by PASS) can be used to identify anomalies in program behavior, which often signal a security intrusion or exploit. Using provenance-based histograms, we can construct models of the regular usage patterns for a particular program. When the intrusion or exploit results in behavior deviating from these models, its provenance allows us to flag it as a potential intrusion. Replaced binaries that acted significantly differently were especially easy to detect. Histograms on simple provenance statistics are good for identifying simple anomalies, but more complex statistics (see future work) may enable us to detect subtler intrusions.
